Compositional equivalences based on Open pNets

### 2020-02-10 2016-Weak-FH-Bisim/WeakBisim.tex
comments figures

Fig. \ref{SimpleProt:Spec} (left) shows the example principle: two unspecified
processes $P$ and $Q$ (holes) communicate messages, with a data value
argument, through the two protocol entities. Process $P$ send an $p\_send(m)$ message to the $Sender$; this communication is denoted as $\underline{in(m)}$, where underline actions represent internal actions (similar to the $\tau$ of CCS), that cannot be further synchronised with the environment. At the other end process $Q$ receives the message from the $Receiver$. The holes $P$ and $Q$ can also have other interactions with their environment, represented here by actions $p\_a$ and $q\_b$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect transport errors, and signal them to the sender. The return $ack$ message from $receiver$ to $sender$ is supposed to be safe. The final transmission of the message to the recipient (the hole Q) includes the value of the ``error counter'' $ec$.

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet of the system \emph{specification}:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystemSpec} is the top level of the pNet structure. It acts as the parallel operator. It consists of tree nodes: two Holes $P$ and $Q$ and one subnet, noted {\it PrefectBuffer}. Nodes of the tree are synchronized using four synchronization vectors, that  express the possible synchronization between the parameterized actions of a subset of the nodes. For instance, in the vector "$< p\_send(m),in(m),\_> \rightarrow \underline{in(m)}$ only $P$ and {\it PrefectBuffer} nodes are involved in the synchronization. More precisely, the synchronization between these processes occurs when  process $P$  performs $p\_send(m)$  action sending a message, and  the medium accepts the message through $in(m)$ action  at the same time. And the result that will be returned at upper level is the internal action  $\underline{in(m)}$.

Fig. \ref{SimpleProt:Impl} shows the pNet model of the protocol implementation. When the $Medium$ detects an error (modeled by a local $\tau$ action), it sends back a $m\_error$ message to the $Sender$. The $Sender$ increments its local error counter $ec$, and resends the message (including $ec$) to the $Medium$, that will, eventually, transmit $m,ec$ to the $Receiver$.

@@

\begin{figure}[ht]
   \centerline{\includegraphics[width=9cm]{XFIG/SPSpecOpen}}
   \caption{Specification OA}
   \label{SimpleProtCounter:SpecOA}
\end{figure}

\paragraph{Example} Figure \ref{SimpleProtCounter:SpecOA} shows the open automaton computed from the Specification pNet in Figure \ref{SimpleProt:Spec}. 
\RAB{Here, in the open automaton drawing, we use a user-friendly notation for the OTs, in the form ``$\{actions of holes\} [guard] action \{assigns\}$'', rather than the full OT form with the red-dotted rule. Note that the resulting action of the OTs appear here between the guard and the set of assignments. For example in the upper left of the figure, the OT where $\beta=\{P\mapsto p\_a\}$, $\Pred=[\forall m. P\_a\neq p\_send(m)]$, and $\alpha=P\_a$ is denoted as ``\texttt{\{P->p-a\} [Forall m.P-a!=p-send(m)] P-a}''.
}

    \begin{figure}[ht]
  %\centerline{\includegraphics[width=10cm]{ATG/SPImplOpen}}
  \caption{Open Automaton of the Simple Protocol Implementation}  %\label{SimpleProtCounter:ImplOA}
\end{figure}

    \begin{figure}[ht]
  \centerline{\includegraphics[width=10cm]{XFIG/SPImplOpen}}
  \caption{Open Automaton of the Simple Protocol Implementation}  \label{SimpleProtCounter:ImplOA}
\end{figure}

    Figure \ref{SimpleProtCounter:ImplOA} shows the open automaton of the implementation $SimpleSystem$.
    \ERIC{In this drawing, we have short labels for states, representing $\ostate{s_0,m_0,r_0}$ by \texttt{000}. For readability we only have written the \emph{external} actions of holes $P$ and $Q$ on the loop transitions of the first staes; they are also present on all other states. 
The resulting behavior is quite simple:  we have a main loop including receiving a message from $P$ and transmitting the same message to $Q$, with some intermediate $\tau$ actions from the internal communications between the protocol processes. In most of the transitions, you can observe that data is propagated between the successive state variables (holding the message, and the error counter value). On the right of the figure, you have a loop of $\tau$ transitions showing the handling of errors and corresponding increments of the error counter.}

@@

\begin{figure}[ht]
   \centerline{\includegraphics[width=7cm]{XFIG/WOT2-result}}
  \caption{Construction of WOT2}
   \label{WOT2}
\end{figure}

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.

@@

         For any of these, apply again WT3, using the central \texttt{q\_recv(ec)} transition as $\alpha$, and you get:
      $\openrule
         {\{Q\mapsto\texttt{q\_revc(ec)}\subst{ec\gets ec+n}\}, True,
			\subst{ec\gets ec+n} }
         {s \OTWeakarrow {\texttt{out(m,ec)}\subst{ec\gets ec+n}} s'}$
\end{example}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=13cm]{ATG/SPSpecWeakOpen}}
  \caption{Weak Open Automaton of the Specification \TODO{out of date}}
   \label{SimpleProtCounter:WeakSpecOA}
\end{figure}

>>>>

Fig. \ref{SimpleProt:Spec} (left) shows the example principle: two unspecified
processes $P$ and $Q$ (holes) communicate messages, with a data value
argument, through the two protocol entities. Process $P$ send an $\texttt{p\_send(m)}$ message to the $Sender$; this communication is denoted as $\texttt{\underline{in(m)}}$, where underline actions represent internal actions (similar to the $\tau$ of CCS), that cannot be further synchronised with the environment. At the other end process $Q$ receives the message from the $Receiver$. The holes $P$ and $Q$ can also have other interactions with their environment, represented here by actions $\texttt{p\_a}$ and $\texttt{q\_b}$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect transport errors, and signal them to the sender. The return $ack$ message from $receiver$ to $sender$ is supposed to be safe. The final transmission of the message to the recipient (the hole Q) includes the value of the ``error counter'' $ec$.

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet of the system \emph{specification}:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystemSpec} is the top level of the pNet structure. It acts as the parallel operator. It consists of tree nodes: two Holes $P$ and $Q$ and one subnet, noted {\it PrefectBuffer}. Nodes of the tree are synchronized using four synchronization vectors, that  express the possible synchronization between the parameterized actions of a subset of the nodes. For instance, in the vector "$< \texttt{p\_send(m)},\texttt{in(m)},\_> \rightarrow \texttt{\underline{in(m)}}$ only $P$ and {\it PrefectBuffer} nodes are involved in the synchronization. More precisely, the synchronization between these processes occurs when  process $P$  performs $p\_send(m)$  action sending a message, and  the medium accepts the message through $\texttt{\texttt{in(m)}}$ action  at the same time. And the result that will be returned at upper level is the internal action  $\texttt{\underline{in(m)}}$.

Fig. \ref{SimpleProt:Impl} shows the pNet model of the protocol implementation. When the $Medium$ detects an error (modeled by a local $\tau$ action), it sends back a $\texttt{m\_error}$ message to the $Sender$. The $Sender$ increments its local error counter $ec$, and resends the message (including $ec$) to the $Medium$, that will, eventually, transmit $m,ec$ to the $Receiver$.

@@

\begin{figure}[ht]
   \centerline{\includegraphics[width=12cm]{XFIG/SPSpecOpen}}
   \caption{Specification OA}
   \label{SimpleProtCounter:SpecOA}
\end{figure}

\paragraph{Example} Figure \ref{SimpleProtCounter:SpecOA} shows the open automaton computed from the Specification pNet given in Figure \ref{SimpleProt:Spec}.

%    \begin{figure}[ht]
  %\centerline{\includegraphics[width=10cm]{ATG/SPImplOpen}}
%  \caption{Open Automaton of the Simple Protocol Implementation}  %\label{SimpleProtCounter:ImplOA}
%\end{figure}

 \begin{figure}[ht]
  \centerline{\includegraphics[width=15cm]{XFIG/SPImplOpen}}
  \caption{Open Automaton of the Simple Protocol Implementation}  \label{SimpleProtCounter:ImplOA}
\end{figure}

    Figure \ref{SimpleProtCounter:ImplOA} shows the open automaton of the implementation $SimpleSystem$ (Figure \ref{SimpleProt:Impl}).
    \RAB{In this drawing, we have short labels for states, representing $\ostate{s_0,m_0,r_0}$ by \texttt{000}. Note that open transitions are denoted $\texttt{OT}_i$ and tau open transition by  $\texttt{OT}_tau$. The resulting behavior is quite simple:  we have a main loop including receiving a message from $P$ and transmitting the same message to $Q$, with some intermediate $\tau$ actions from the internal communications between the protocol processes. In most of the transitions, you can observe that data is propagated between the successive state variables (holding the message, and the error counter value). On the right of the figure, you have a loop of $\tau$ actions showing the handling of errors and corresponding increments of the error counter.}

    \ERIC{In this drawing, we have short labels for states, representing $\ostate{s_0,m_0,r_0}$ by \texttt{000}. For readability we only have written the \emph{external} actions of holes $P$ and $Q$ on the loop transitions of the first staes; they are also present on all other states. 
The resulting behavior is quite simple:  we have a main loop including receiving a message from $P$ and transmitting the same message to $Q$, with some intermediate $\tau$ actions from the internal communications between the protocol processes. In most of the transitions, you can observe that data is propagated between the successive state variables (holding the message, and the error counter value). On the right of the figure, you have a loop of $\tau$ transitions showing the handling of errors and corresponding increments of the error counter.}

@@

\begin{figure}[ht]
   \centerline{\includegraphics[width=8cm]{XFIG/WOT2-result}}
  \caption{Construction of WOT2}
   \label{WOT2}
\end{figure}

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.
\RAB{The weak open transition here is the transition delivering the result of the algorithm to hole $Q$ by applying or rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $\texttt{WT}_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$;  $\openrule
         {\{\}, True,
			() }
         {b1 \OTWeakarrow \tau b1}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {b1 \OTWeakarrow \tau b1}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {b1 \OTWeakarrow \tau b1} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{q\_recv(ec)}  as $\alpha$, we get:
      $\openrule
         {\{Q\mapsto\texttt{q\_revc(ec)}\subst{ec\gets ec+n}\}, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {b1 \OTWeakarrow {\texttt{out(m,ec)}\subst{ec\gets ec+n}} b0} \forall n \geq 0$. }

@@

         For any of these, apply again WT3, using the central \texttt{q\_recv(ec)} transition as $\alpha$, and you get:
      $\openrule
         {\{Q\mapsto\texttt{q\_revc(ec)}\subst{ec\gets ec+n}\}, True,
			\subst{ec\gets ec+n} }
         {s \OTWeakarrow {\texttt{out(m,ec)}\subst{ec\gets ec+n}} s'}$

@@

%\begin{figure}[h]
%   \centerline{\includegraphics[width=13cm]{ATG/SPSpecWeakOpen}}
%  \caption{Weak Open Automaton of the Specification \TODO{out of date}}
%   \label{SimpleProtCounter:WeakSpecOA}
%\end{figure}
